### 2021-03-15 analysis/paper/paper.Rmd
added basic plot of data

---
title: "Correlations between gender and research topics at three major archaeology conferences"
author:
  - Yichun Chen:
      email: yichun33@uw.edu
      institute: [UofO]
      correspondence: true
  - Ben Marwick:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

# Abstract

Disproportionate representation of genders has long existed in many academic fields. Rising interest in gender equality in society generally has resulted in greater scrutiny on gender inequality in academic communities. Analysis of authorship of peer-reviewed publications shows that archaeology is similar to other academic fields in having long been dominated by men. We ask if gender disproportionality is evident in the choice of topics that archaeologists present on at major conferences, particularly the Society of American Archaeology (SAA), the European Association of Archaeologists (EAA) and the Computer Applications and Quantitative Methods in Archaeology (CAA) meetings. Does the gender of the participants in these archaeological conferences correlate with the topics of their presentations? We analysed presenters' names in the published programs of these three archaeology conferences to infer gender. We then used machine learning to identify topics from presentation titles. We found that there are some associations between gender and topics. An awareness of these correlations between research topics and gender is important to ensure equitable participation in archaeology, and unbiased access to training opportunities for students. We expect these findings to be useful for instructors who prioritize gender equality in student and early career research activities.

# Introduction

Gender under-representation and inequality has long existed in science, from under-representation of women and minority genders in school text books to academic publications [@hamilton2006gender; @tushingham2017peer]. Women are often under-represented for several reasons, including stereotyping that unfairly underestimates individuals' abilities, which often leads to lack of support for women in academic fields [@xu2008gender]. In this paper we explore the relationship between gender and research topics to determine if women and men tend to focus more on certain topics than others.

We used a computational method to identify topics from the titles of presentations delivered at meetings of the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017). We analysed the covariance of topics and genders of presenters to explore how gender ratios vary between these conferences and by topics.

# Background

According to the American Anthropological Association's (AAA) _AnthroGuide_, there is a growing number of women participating in anthropology's graduate programs. For example, in the 2012-2013 school year, women consists of 64% of graduate students in anthropology programs and 65% of students that completed an anthropology doctoral degrees are women, whereas in 1972 32% of PhD are granted to women and 59% in 1995 [@bardolph2014critical]. Although women slightly outnumber men in anthropology graduate programs, previous work has shown substantial gender imbalances in publication practices in archaeology where men greatly outnumber women, especially as first authors in peer reviewed journal articles [@bardolph2014critical]. One of the first studies of this imbalance is @bardolph2014critical, who looked at 7,381 members of the Society of American Archaeology and found 53% men and 47% women. She also looked at 1,601 articles published during 1990-2013 in five high-visibility archaeology journals and found a range of proportions of women to men authors of 0.32-0.73, with three of the five journals having a proportion of <0.5.

@rodriguez2018gender studied gender balance among 309 authors of 299 publications produced during 1978-2016 by the Atapuerca Project, a large archaeo-palaeontological project in Burgos, Spain. They 'manually defined' the gender of each author, and found 13 articles authored exclusively by women, 34.8% of papers have women as first-authors, and overall, 35.9% of all authors are women [@rodriguez2018gender]. Over time, they observed a trend of increasing numbers of papers that include women as authors, numbers of paper lead by women as the first author, and increases in the ratio of female to male authors in group-authored papers. They did not make any observations about relationships between gender and topics.

Another similar imbalance can be seen in terms of the authorship of articles or papers written. For example, for the authorship of the first authors of 1,104 articles published in the _Journal of Field Archaeology_ during 1974–2018, 72% of them are men [@heath2020guest]. @heath2020guest investigates the possibility that the gender gap in authorship is due to due to sexism in the peer review process. Out of 830 instances of peer review for where both the reviewer’s gender and the first author's gender could be determined, neither the first author’s gender, the reviewer’s gender, nor the combination of the two had a significant effect on the reviewer's recommendation.

@fulkerson2019dominates collected data on author's gender and occupational affiliation in peer reviewed journals (_American Antiquity_ (AQ), _Advances in Archaeological Practice_ (AAP), _Journal of California and Great Basin Anthropology_ (JCGBA), and _California Archaeology_ (CA)) and some non-peer-reviewed venues (the _SAA Archaeological Record_ (SAA Record) and the SCA _Proceedings_). Among the 5,010 authors of 2,445 articles in their sample, 27.1% of first/single authors of peer-reviewed journal articles are women, and 72.9% are men. The gender gap is less pronounced in the non-peer-reviewed venues with women accounting for 40.8% of lead authors in 517 articles.

@bardolph2018controlling examined data from 2007 to 2017 about the membership in the Society of California Archaeology (SCA) conference, and the lead-authors of JCGBA and CA. @bardolph2018controlling noted that women's conference presentation rates are consistent with their membership rates in SCA, and that it was not until 2017 when women's participation rates finally exceeded men's. Although women are more actively participating in SCA, this is not the case for lead-authorship of JCGBA and CA [@bardolph2018controlling]. Throughout the majority of time in the period studied, lead-authorship of JCGBA articles is highly skewed towards men with only 34% women. A greater difference between men and women lead-authorship can be seen in CA, in 2009 there was no women lead-authors at all, and overall only 23% of the published papers of CA have women as lead-authors [@bardolph2018controlling].

When looking at gender ratios among the membership in Society for California Archaeology from 1967 to 2016, @tushingham2017peer report a trend of women increasingly maintaining their society membership they remain underrepresented in peer-reviewed journals. @tushingham2017peer examined authorship gender trends in 1,599 papers in three journals, _Journal of California and Great Basin Anthropology/Journal of California Archaeology_ (JCGBA/JCA), _California Archaeology_ (CA) and a non-peer reviewed _Proceedings_ of the Society for California Archaeology (PSCA). In a total of 2,617 authors, 844 (32.3%) were women, 1,762 (67.3%) were men, and 11 (0.4%) were gender unknown/ambiguous [@tushingham2017peer]. They found significant increases in the proportion of female lead authors over time in JCGBA/JCA and PSCA, but not in CA, perhaps because of the shorter publishing period for this journal [@tushingham2017peer]. This similar trend of a gender gap in archaeology can also be seen in women's pay [@vanderwarker2018ucsb]. In the University of California, Santa Barbara Gender Equity Project,@vanderwarker2018uscb report an imbalance in pay rates, for example, women make a greater proportion of people paid less \$60,000 and men outnumber women in salary brackets greater than \$60,000. [@vanderwarker2018uscb].

Studies of authorship where that gender is inferred from the author's first name are common, but can give wrong results for non-English, androgynous, and uncommon names. @heathwrites avoided these limitations by conducting a survey that directly asked archaeologists for their self-identifications of gender, race/ethnicity and sexual orientation. In her intersectional study, @heathwrites a higher number of women publishing in archaeology than shown in previous studies, even in the studies of the same publications that she studied. @heathwrites explains this as a result of change over time, with her study including more recent publications than others, and concludes that most journals is slowly reaching gender parity.

@sinclair2016intellectual, compared the relative standings between men and women researchers in terms of number of citations made by other authors in the field of archaeology. Among the list of 50 most cited authors, only 6 of them (12%) are women, meaning that recognition by citation is still mostly given to men. A similar result can be seen if we increase the sample to the top 250 authors, with 19% women [@sinclair2016intellectual]. After analyzing the most cited authors and their articles' thematic clusterings, we see that there are relatively more women authors in the category of Archaeology of Early State Societies in the Americas (N and C America) with 46% women authors. Other categories with high proportions of women are Archaeological chemistry (esp. lipids analysis) with 40% women authors, Isotope analysis with 30% women authors, and dating (esp. thermoluminescence and magnetics) with 30% women authors. The average women author cited in the field of archaeology as a whole is around 14%. Sinclair concludes that the higher representation of women as cited authors on topics of laboratory analysis supports Gero's claim that women in archaeology were most active in laboratory-based activities rather than excavation/fieldwork related activities.

We have extended the study of ratios of authorships in the field of archaeology by specifically focusing on the gender imbalance that occurs in the topics that the authors present in major archaeological conferences.

```{r}
library(here)
# column names should be:
# session, FirstName, AbstractTitle

SAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 SAA Title.xlsx"))
EAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 EAA.xlsx")) 
CAA18 <- readxl::read_excel(here::here("analysis/data/raw-data/2018 CAA.xlsx"))

SAA18_row <- nrow(SAA18)      
EAA18_row <- nrow(EAA18)      
CAA18_row <- nrow(CAA18)

library(gender)
lynne_gender <-  gender("Lynne")
```

# Methods

We requested from conference organisers spreadsheets files of the publicly available program information for three major archaeology meetings held in 2018: the Society of American Archaeology (SAA, 2016-2019), the European Association of Archaeologists (EAA, 2018) and the Computer Applications and Quantitative Methods in Archaeology (CAA, 2017).

We estimated the gender of the first-named speaker for each presentation using the R programming language and the gender package [@lincolngender; @mihaljevic2019reflections; @blevins2015jane]. We inferred the gender of personal names by looking each name up in the US Social Security Administration (SSA) baby name data, and calculating the overall probability that a given name was male or female [@mihaljevic2019reflections]. If the proportion of people in the SSA data with a certain given name are recorded as female is 0.5 or higher, we predicted the gender of the presenter with this name as female. For example, in the SSA dataset the the name "Lynne" returns `r lynne_gender$proportion_male` as proportion of individuals with this name who are recorded as male, and `r lynne_gender$proportion_female` as the proportion female, thus we inferred that people in our conference data who are called Lynne are women.

This method of inferring gender has the advantage of speed, transparency and reproducibility, but also some substantial limitations that are important to be upfront about. We are only able to infer binary male/female genders and assign the first names into these two categories. This has the unfortunate result of excluding or mis-identifying other genders from the results, excluding them from our analysis. We considered it impractical and invasive to write to each first-named presented to request their gender information. A further limitation of our approach is that it sometimes fails to classify non-English names at all, as the SSA data consists mostly of English names. This means that people with non-English names are underrepresented in our results. It is important to note that the inferences of gender presented here are not self-identified by the presenters, but are computed probabilistically. Better quality and more representative data would result if presenters self-identified their gender to conference organisers, but currently these data are not available. We encourage conference organizers to collect gender data directly from participants to improve the representation of minority genders in future studies. Our hope is that this work, despite its many limitations, will stimulate the collection of more reliable, justifiable, and useful data by conference organisers.

We identified the topic of each presentation by generating a structural topic model for all the presentations in each of the three conferences. Topic modeling is a machine learning method to automatically find related groups of words that resemble traditional 'themes' or 'topics' in a collection of documents. Topic models are 'unsupervised' methods because they infer rather than assume the content of the topics in a collection of documents, and they have been used across a variety of fields [@chang2009reading]. This approach uses latent Dirichlet allocation to allow every word to be present in every topic, but with different weightings, such that the most heavily weighted 5-10 words of a topic often capture the essence of the topic as a coherent and familiar concept. When generating topic models, we have to first decide on the number of topics that the method will identify in our texts.

We have used the built in algorithm of the stm package based off of Lee and Mimno (2014) to find the number of topics [@mimno2014low]. This method finds the number of topic by choosing anchor words and using statistical probability so that for each assumed topic a specific anchor word will only appear once, after this the algorithm will reconstruct the word co-occurance pattern for the non-anchor words to create a convex combination of the co-occurance patterns of those anchor words. After creating the convex the algorithm will then choose the farthest point from our subspace until the given number of anchors have been found. It is important to note that this method will not alway result in the same number of suggested topics, as it is based off of probability and dependent on the quality of the anchor words chosen. The topic modeling algorithm then assigns all words in every document into our set number of topics, and assigns each word a probability based on its per-topic-per-word probability, and each document (i.e. conference presentation title) is assigned a distribution of topics with different weights. To visualise the topic model output we use the 10 words with the highest weighting within each topic as keywords to represent each topic. In generating the topic model we excluded words that are common in English generally or in archaeology and so have little semantic value, for example, 'the', 'a', 'el', 'al', 'archaeology', etc.

A structural topic model (STM) allows us to include, alongside the typical topic modeling process, other document-level metadata to analyse covariance between topics and metadata [@roberts2019package]. Covariates of interest can be included into the prior distributions for document-topic proportions and topic-word distributions using a standard regression model [@roberts2014structural]. This means that we can examine the relationship between the topics and document-level variables of interest. In our case, we computed the relationship between the inferred the gender of the first-named presenter for each presentation, and the distribution of topics in each presentation. When we observed the topics for each presentation, we generated a regression where the topic is the outcome variable, and gender is the explanatory variable. This regression gives us insights into whether gender caused archaeologists to spend a larger portion of their presentation on a particular topic.

# Results

```{r, eval=FALSE}

# make the custom function available here in this Rmd file
source(here::here("analysis/paper/functions-for-gender.R"))

# compute gender model, take a few minutes...
SAA18_gm <- get_the_gender_data(SAA18)
saveRDS(SAA18_gm, here::here("analysis/data/derived-data/SAA18_gender_model.rds")) # ok

EAA18_gm <- get_the_gender_data(EAA18, first_name_appears_first = FALSE)
saveRDS(EAA18_gm, here::here("analysis/data/derived-data/EAA18_gender_model.rds"))

CAA18_gm <- get_the_gender_data(CAA18)
saveRDS(CAA18_gm, here::here("analysis/data/derived-data/CAA18_gender_model.rds"))
```

```{r, eval=FALSE}

# make the custom function available here in this Rmd file
source(here::here("analysis/paper/function-for-structural-topic-model.R"))

SAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))

EAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))

CAA18_gm <- 
  readr::read_rds(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))

# compute and save the topic model for each conf, takes quite a long time
SAA18_stm <- 
  stm_model_and_plot(conf_paper_titles = SAA18, 
                     conf_gender_results = SAA18_gm, 
                     max.em.its = 1000)

saveRDS(SAA18_stm, here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))

EAA18_stm <- 
  stm_model_and_plot(conf_paper_titles = EAA18, 
                     conf_gender_results = EAA18_gm, 
                     first_name_appears_first = FALSE,
                     max.em.its = 1000)

saveRDS(EAA18_stm, here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))

CAA18_stm <-   
  stm_model_and_plot(conf_paper_titles = CAA18, 
                     conf_gender_results = CAA18_gm, 
                     lower_thresh = 3,
                     max.em.its = 1000)

saveRDS(CAA18_stm, here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))

```

```{r}
# get plots
library(tidyverse)

# SAA -----
SAA18_stm_plot <- 
  readr::read_rds(here::here("analysis/data/derived-data/SAA18_topic_model_and_plot.rds"))

SAA18_stm_plot$plot
ggplot2::ggsave(here::here("analysis/figures/SAA18_stm_plot.png"), 
       height = 7, 
       width = 6)

SAA_stm_topic_number <-
  SAA18_stm_plot$model$settings$dim$K
SAA_stm_topic_number_significant_covary <- 
  nrow(layer_data(SAA18_stm_plot$plot, 3))

# CAA -----
CAA18_stm_plot <- 
  readr::read_rds(here::here("analysis/data/derived-data/CAA18_topic_model_and_plot.rds"))

CAA18_stm_plot$plot
ggsave(here::here("analysis/figures/CAA18_stm_plot.png"), height = 7, width = 6)

CAA_stm_topic_number <-
  CAA18_stm_plot$model$settings$dim$K
CAA_stm_topic_number_significant_covary <- 
  nrow(layer_data(CAA18_stm_plot$plot, 3))

# EAA -----
EAA18_stm_plot <- 
  readr::read_rds(here::here("analysis/data/derived-data/EAA18_topic_model_and_plot.rds"))

EAA18_stm_plot$plot
ggsave(here::here("analysis/figures/EAA18_stm_plot.png"), height = 7, width = 6)

EAA_stm_topic_number <-
  EAA18_stm_plot$model$settings$dim$K
EAA_stm_topic_number_significant_covary <- 
  nrow(layer_data(EAA18_stm_plot$plot, 3))
```

```{r, eval=TRUE}
SAA18_gender <-
  readRDS(here::here("analysis/data/derived-data/SAA18_gender_model.rds"))
EAA18_gender <-
  readRDS(here::here("analysis/data/derived-data/EAA18_gender_model.rds"))
CAA18_gender <-
  readRDS(here::here("analysis/data/derived-data/CAA18_gender_model.rds"))

count_gendered_SAA18 <- nrow(SAA18_gender)
perc_gendered_SAA18 <- round(nrow(SAA18_gender) / SAA18_row, 2) * 100
female_count_SAA18 <- 
  length(which(SAA18_gender$gender == "female"))
male_count_SAA18 <- 
  length(which(SAA18_gender$gender == "male"))
female_to_male_SAA18 <- 
  round(female_count_SAA18 / male_count_SAA18,1)
missing_gender_SAA18 <- 
  SAA18_row - (female_count_SAA18 + male_count_SAA18)

count_gendered_EAA18 <- nrow(EAA18_gender)
perc_gendered_EAA18 <- round(nrow(EAA18_gender) / EAA18_row, 2) * 100
female_count_EAA18 <- 
  length(which(EAA18_gender$gender == "female"))
male_count_EAA18 <- 
  length(which(EAA18_gender$gender == "male"))
female_to_male_EAA18 <- 
  round(female_count_EAA18 / male_count_EAA18, 1)
missing_gender_EAA18 <-
  EAA18_row - (female_count_EAA18 + male_count_EAA18)

count_gendered_CAA18 <- nrow(CAA18_gender)

perc_gendered_CAA18 <- round(nrow(CAA18_gender) / CAA18_row, 2) * 100

female_count_CAA18 <- 
  length(which(CAA18_gender$gender == "female"))

male_count_CAA18 <- 
  length(which(CAA18_gender$gender == "male"))

female_to_male_CAA18 <- 
  round(female_count_CAA18 / male_count_CAA18, 1)

missing_gender_CAA18 <-
  CAA18_row - (female_count_CAA18 + male_count_CAA18)
```

```{r}
# some basic visualisations to show the data
list_confs <- 
  list(SAA18_gender %>% 
              mutate(session = as.character(session)),
       CAA18_gender,
       EAA18_gender %>% 
              mutate(session = as.character(session))
       )

names(list_confs) <- c("SAA", "CAA", "EAA")

all_conf_tbl <- 
  bind_rows(list_confs,
            .id = "conf")

# how many papers per conf?
ggplot(all_conf_tbl) +
  aes(conf,
      fill = gender) +
  geom_bar(position = "dodge") +
  theme_minimal(base_size = 16) +
  theme(legend.position = c(.2,.60)) +
  labs(x = "", 
       y = "number of papers presented in 2018")
```

Overall we see similar women-to-men ratios for the SAA (`r female_to_male_SAA18`) and EAA (`r female_to_male_EAA18`), but much lower for the CAA (`r female_to_male_CAA18`). Even before we look into the gender preferences for specific topics within each conference, we already have a hint that computational topics may be less preferred by women than by men.

In the SAA data we found a total of `r prettyNum(SAA18_row, big.mark = ",")` presentations. Of these we could identify the first-named authors of `r count_gendered_SAA18` (`r perc_gendered_SAA18`%) presentations as either men (n=`r male_count_SAA18`) or women (n=`r female_count_SAA18`). In the SAA presentations we identified the optimum number of topics as `r SAA_stm_topic_number`, with `r SAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter. Eight of these topics were associated with men, and nine topics associated with women. One of the main difference that we can see between the topics of the genders are the locations that the presenters are interested in.  Topics associated more with women include bioarchaeology, cemeteries, burials, shells, and ceramics. Topics associated more with men include survey and landscape archaeology, geoarchaeology, rituals, and regional studies. Women are more likely to work on locations such as Mesoamerica, Arizona and national parks, while men are more likely to present on topics about the Great Lakes, and Honduras.

For the EAA data we have `r prettyNum(EAA18_row, big.mark = ",")` presentations, and `r count_gendered_EAA18` (`r perc_gendered_EAA18`%) where the first author could be classified as either a man (n=`r male_count_EAA18`) or woman (n=`r female_count_EAA18`). For EAA we generated a total of `r EAA_stm_topic_number` topics, with `r EAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter. Seven of these topics were associated with men, and nine topics associated with women. Although topics significantly associated with men and women both include the word "Mediterranean", women seem to be focusing more on the western Mediterranean while men seems to be focusing on the northern Mediterranean. Topics associated more with women include burials, graves, museums, animals, diet, and ritual. Topics associated more with men include geoarchaeology, architecture, and towns.

For the CAA meeting we have `r prettyNum(CAA18_row, big.mark = ",")` presentations, and `r count_gendered_CAA18` (`r perc_gendered_CAA18`%) first-named authors could be classified as either men (n=`r male_count_CAA18`) or women (n=`r female_count_CAA18`). For CAA we generated a total of `r CAA_stm_topic_number` topics, with `r CAA_stm_topic_number_significant_covary` of these showing non-random co-variance with the gender of the first-named presenter. Three of these topics were associated with men, and three topics associated with women. Topics associated more with women include archives, GIS, geophysics, collaboration, and teaching. Topics associated more with men include environmental and landscape archaeology.

# Discussion

Our results show significant correlations between the gender of the first-named presenter and the topics in their conference presentation in all three conferences. Although each of the three meetings have distinctive sets of topics, we can identify some common themes in the topics that correlate with gender. We found that topics strongly associated with women at all three conferences relate to managing cultural heritage, GIS, and isotope analyses. Pair-wise shared topics between conferences also reveal the norms of each community. These contrasts and similarities reveal some of the choices that presenters make when deciding where to share their work. The SAA and EAA are defined mostly by the geographic region of their community, the Americas and Europe. They have more participants, and our model identified a higher number of topics in these two, compared to CAA. CAA, with its focus on computer applications and quantitative methods, has a much smaller community, and a small number of topics. So we might expect less in common between CAA and the two bigger conferences. At both the SAA and EAA we see that topics about burials, cemeteries, bodies, and graves are strongly associated with women. For the SAA and CAA the shared women-associated topics include learning and practice. For the EAA & CAA the common topics associated with women presenters are self-referential: women and gender. The absence of bioarchaeology and topics about human remains at CAA suggest that either archaeologists working on those topics do not recognise this conference as a meaningful place to present that work, or that there are some unrealised opportunities to apply computational and quantitative methods in bioarchaeological research.

Topics that are strongly associated with men at all three conferences include geoarchaeology and geophysics. Topics associated with men at the SAA and CAA include built-location-based research, indicated by keywords such as house/town/site/fort. At both SAA and CAA we see landscape as a shared topic that is strongly associated with men. For the EAA and CAA there is little overlap in topics strongly associated with men. The geophysics topic is notable here because it is also in the CAA topic that is most strongly associated with women. It is strongly associated with men at the SAA and EAA, but also strongly associated with women at the CAA. This indicates that a topic is not immutably bound to either men or women, but that topics can shift in prominence among genders, depending on the context. This is important for understanding the relationship between genders, topics, and communities. These relationships are flexible and contingent, such that a topic-gender association is specific to a particular community (e.g. conference), and argues against topic-gender generalisations that attempt to transcend communities of practice.

Related work on gender differences in research topics has noted major differences in the proportions of men and women across many academic-related interests. One widely discussed theory, especially in psychology, to explain this is the theory of people-thing interest dimensions which proposes that men have stronger interest in things and their mechanisms, while women have a stronger interest in persons and their feelings [@su2009men]. However, recent work by @thelwall2019gender shows that gender differences in choice of research topics cannot be fully explained by people-thing theory from psychology. Thewall et al. looked at 508,283 journal articles, classified the articles into research areas, and determined the authors' genders by matching names with the 1990 US census data. Instead of a people-thing contrast, they find that women more likely to use exploratory and qualitative methods rather than quantitative methods, and men show more interest in male in abstraction, and power/control fields.  In our data, we can similarly see in an interest from men in quantitative methods. However, our results do not so clearly divide the interests of the two genders we have data on, for example, with women are more prominent in GIS, and isotope analyses, which are highly quantitative areas.

@ostapenko2018textual identified common topics in the personal statements of aspiring surgeons that were specific to men and women authors. Women tended to discuss surgery as working as a team, while men focused on their specific individual clinical experiences. @ostapenko2018textual propose that the differences between male and female statements may come from actual motivating factors for career goals and aspirations, or they may reflect differences in beliefs about what makes a successful personal statement. They could not evaluate which explanation was most important, but a key observation is that authors of the personal statements may not deliberately choose the themes for their statements. Instead, they may be using themes that they have received positive feedback on on the past, that has perhaps unconsciously biased them towards certain themes in their writing. Similarly archaeologists might choose topics for their research based on a pattern of positive or negative feedback on topic choices over time, and so their choices are not fully deliberate.

Theories such as people-thing and Thewall et al's exploratory/abstraction assume that researchers freely make choices about their research preferences, and that these choices are influenced by inborn cognitive differences. However, these approaches do not fully explain gender differences in topic choices. A meta-analysis of gender and science research by the European Commission [@caprile2012meta] found that structural and life-course factors also profoundly shape the distribution of genders in research fields. Socialisation factors, such as media representation of scientists and family role models strongly shape career choices. Life-course factors include the 'rush hour' which is the time when family and academic demands collide and pivotal decisions are made about having children and establishing an academic career. Historically this is a time that leads to women making different choices to their male peers, but this can also affect caregivers of any gender. For example, caregivers may be less mobile to conduct field research or relocate for career advancement. These structural and life-course factors may add up to subtle, but pervasive exclusionary practices the skew gender representation in research fields that are independent of individual choices and preferences.

A key limitation of this study is our concept of gender, and how we measure it. Like many social science studies, we have measured gender assuming it has two binary categories. However, we recognise that many cultures have long included more than two genders [@wilson1996we; @graham2004s], and there is a growing awareness of this in Western cultures also. A better measure of gender would represent it as a multifaceted spectrum [@tobin2010intrapsychics]. The problem with our binary measure is that it does not reflect our current understanding of gender, forces people in misclassified categories, and is hostile to public acceptance and advocacy for transgender and nonbinary individuals. Our instrument for collecting gender data only returns two categories, so this has limited potential for overcoming this limitation. Instead, we recommend that conference organisers collect gender information directly from presenters using inclusive gender measures. For example, by providing nonbinary options for presenters on registration forms, and by asking about gender as an open‐ended question for participants to self-identify. We recommend that future work on gender in archaeology avoid 'othering' language when describing results. For example, reporting a sample as '150 participants (48% women; 49% men; 3% other)' violates ethical standards because such wording implies that binary gender is normal or appropriate, whereas trans and nonbinary gender is not (it is 'other'). Our hope is that our study might be the last one to  discuss gender as a binary.

# Conclusion

Although the gender gap in academia has been shrinking in recent years, there remain substantial differences, which often result in underrepresentation of women in many fields and career stages, as well as disadvantages such as lower salaries and less access to high-prestige employment [@caprile2012meta]. Even research into these gender biases suffers a bias, with @Cislak_Formanowicz_Saguy_2018 finding that articles on gender bias and race bias are funded less often and published in journals with a lower Impact Factor than articles on comparable instances of social discrimination.

how to use these results... implicit bias in assigment topics...

# Acknowledgements

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:
u
```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```